Prediction of 2024 US election ...

### 2024-11-03 paper/paper.qmd
changed metholodology chart to boxplot

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  filter(!is.na(methodology)) %>% 
  group_by(methodology, candidate_name) %>%
  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(mapping = aes(x = avg_pct, y = methodology, fill = candidate_name)) + 
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    x = "Average Polled Support (%)", 
    y = "Methodology",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

```
Figure @fig-avg-pct-methodology indicates minimal bias in polling percentages when comparing different polling methodologies, as most polls show an even distribution of support between candidates. However, polls only incorporate IVR (Interactive Voice Response) tend to skew towards Harris

@@

# this model is really here just for explain, this realtionship should not
# be treated as lienar
cor(graph_data$sample_size, graph_data$pollscore)
lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
summary(lm_score_sample)

```
Figure @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll. The figure also shows that most polls have a sample size around 800 to 1200 participants. It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.

@@

```{r}
#| label: tbl-methodology-candidate-support
#| fig-cap: pct vs methodology
#| echo: false
#| warning: false

cleaned_data %>% 
  filter(!is.na(methodology) & !is.na(pct)) %>% 
  group_by(methodology, candidate_name) %>% 
  summarise(methodology, candidate_name, n = n(), avg_pct = mean(pct), sd = sd(pct)) %>% 
  arrange(methodology, candidate_name) %>% unique()

```
Based on @tbl-methodology-candidate-support, we might need to remove methodology from the model.

```{r}
#| label: tbl-state-candidate-support
#| fig-cap: pct vs methodology
#| echo: false
#| warning: false

cleaned_data %>% 
  filter(!is.na(state) & !is.na(pct)) %>% 
  group_by(state, candidate_name) %>% 
  summarise(state, candidate_name, n = n(), avg_pct = mean(pct), sd = sd(pct)) %>% 
  arrange(state, candidate_name) %>% unique()

@@

state_colors <- c(
  alabama = dark_red,
  arizona = purple,
  arkansas = dark_red,
  california = dark_blue,
  colorado = dark_blue,
  connecticut = dark_blue,
  delaware = dark_blue,
  `district of columbia` = dark_blue,
  florida = light_blue,
  georgia = purple,
  idaho = dark_red,
  illinois = dark_blue,
  indiana = dark_red,
  iowa = light_red,
  kansas = dark_red,
  kentucky = dark_red,
  louisiana = dark_red,
  maine = "gray",
  maryland = dark_blue,
  massachusetts = dark_blue,
  michigan = purple,
  minnesota = light_blue,
  mississippi = dark_red,
  missouri = dark_red,
  montana = dark_red,
  nebraska = "gray",
  nevada = purple,
  `new hampshire` = dark_blue,
  `new jersey` = dark_blue,
  `new mexico` = dark_blue,
  `new york` = dark_blue,
  `north carolina` = purple,
  `north dakota` = dark_red,
  ohio = light_red,
  oklahoma = dark_red,
  oregon = dark_blue,
  pennsylvania = purple,
  `rhode island` = dark_blue,
  `south carolina` = dark_red,
  `south dakota` = dark_red,
  tennessee = dark_red,
  texas = light_red,
  utah = dark_red,
  vermont = dark_blue,
  virginia = light_blue,
  washington = dark_blue,
  `west virginia` = dark_red,
  wisconsin = purple,
  wyoming = dark_red
)

>>>>

cleaned_data %>%
  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
  filter(!is.na(methodology)) %>% 
  ggplot(mapping = aes(x = methodology, y = pct)) + 
  geom_boxplot(width = 0.6, position = position_dodge(width = 0.8), outlier.shape = NA) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Methodology",
    y = "Polled Support (%)",
    fill = "Candidate"
  ) +
  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
  theme(legend.position = "bottom") + 
  theme(axis.text.y = element_text(size = 8))

```
Figure @fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.

@@

# this model is really here just for explain, this realtionship should not
# be treated as lienar
# cor(graph_data$sample_size, graph_data$pollscore)
# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
# summary(lm_score_sample)
```
Figure @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll. The figure also shows that most polls have a sample size around 800 to 1200 participants. It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.

@@

@@

state_colors <- c(
  alabama = dark_red,
  arizona = purple,
  arkansas = dark_red,
  california = dark_blue,
  colorado = dark_blue,
  connecticut = dark_blue,
  delaware = dark_blue,
  `district of columbia` = dark_blue,
  florida = light_blue,
  georgia = purple,
  idaho = dark_red,
  illinois = dark_blue,
  indiana = dark_red,
  iowa = light_red,
  kansas = dark_red,
  kentucky = dark_red,
  louisiana = dark_red,
  maine = "gray",
  maryland = dark_blue,
  massachusetts = dark_blue,
  michigan = purple,
  minnesota = light_blue,
  mississippi = dark_red,
  missouri = dark_red,
  montana = dark_red,
  nebraska = "gray",
  nevada = purple,
  `new hampshire` = light_blue,
  `new jersey` = dark_blue,
  `new mexico` = light_blue,
  `new york` = dark_blue,
  `north carolina` = purple,
  `north dakota` = dark_red,
  ohio = light_red,
  oklahoma = dark_red,
  oregon = dark_blue,
  pennsylvania = purple,
  `rhode island` = dark_blue,
  `south carolina` = dark_red,
  `south dakota` = dark_red,
  tennessee = dark_red,
  texas = light_red,
  utah = dark_red,
  vermont = dark_blue,
  virginia = light_blue,
  washington = dark_blue,
  `west virginia` = dark_red,
  wisconsin = purple,
  wyoming = dark_red
)
